master thesis

### 2018-02-24 Paper - Latex/Paper - Skizze - v2.tex
Paper - Results - Results of castor - biological view

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{tag tree}
      \anmerkung{instead of 'tag tee' simulating states and transitions between them (Emanuel)} \\
      At this point we want one fully tagged tree, and one less tagged tree which looks like our real 
        data.\\
      Let's say the first specie (the root node) was free-living (start with a parasite without a host 
        makes no sence). For every transition from a node to his child, we take a random number from
        the father distribution. We decided that from the biological perspective a beta distribution
        reflects our transition probabilities best (see Figure 2.1 \todo{ref einfügen}). \\
      \anmerkung{why is that the case and why is that from a biological perspective? (Bernhard)} \\
      \anmerkungstext{I'd rather say that to ensure that the parameter of the binomial distribution is 
        restricted to the [0,1] interval, we model it... (Bernhard)} \\
      For example when our father node was free-living, then we take from the free-living beta
        distribution. Is the number under the threshold for beeing parasite, we get a change and tag
        the current node as parasite. Otherwise we tag it as free-living. \\
      With this procedure we traverse through the tree from the root to every leaf node. A part of
        this code you see here:
      \begin{lstlisting}[gobble=6]
        from numpy import random
        if father_tag == 0:
          # freeliving_distribution:
          new_random = random.beta(a=A_FL, b=B_FL)
        else:
          # parasite_distribution:
          new_random = random.beta(a=A_P, b=B_P)
        tag = 0         # $\rightarrow$ FL
        if new_random < percentage_parasites:
          tag = 1       # $\rightarrow$ P
      \end{lstlisting}
      \begin{figure}
        \caption{60\% Free-living - 40\% Parasites}
        \centering
          \includegraphics[width=0.5\textwidth]{Figures/40-60.png}
      \end{figure}
      \todo{Bessere Beschriftung, Plot neu erstellen! U.a. mit threshold} \\
      We save each tag with the associated node ID in a nodelist. \\
      \anmerkung{simulationg loss of information (1) states (2) topology/multifurcation (Emanuel)} \\
      The real tree has much less information, we have only information from some current species 
        (leaf nodes) and \todo{and probably negligible internal nodes}. \\
      To simulate our real tree we save for every node an empty placeholde except for some leaf nodes.
      There we save the tags again. The amount of these unknown information is one parameter, which we
      got from our real tree. Or which we can change to \todo{...}
      \todo{Was hiervon gehört ind Methoden, was schon in Implementierung oder ganz woanders hin?}

@@

        Der Fitch Algorithmus ist aufgeteilt in drei Schritte, in welchen man jeweils durch den Baum traversiert.
        Schritt 1 beginnt von de Blättern aus, da sich dort zu Beginn die einzige Information befindet. 
        Für jeden Knoten gilt, wenn seine Kinder schon Information enthalten, dann bilde die Schnittmenge 
        der Tags und schreibe diese als Information in den aktuellen Knoten. Ist die Schnittmenge leer, 
        dann schreibe die Vereinigung aller möglichen Tags in den Knoten. Für alle Kinder, die noch keine 
        Information haben, führe diesen Schritt erst für diese aus.
        Schritt zwei geht von den Kindern der Wurzel bis zu den Vätern der Blätter. Jeder Knoten bekommt 
        einen zweiten Tag, der sich aus der Vereinigung des Tags des Vaterknoten und der Geschwisterknoten
        zusammensetzt. Ist diese leer, bekommt der Knoten wieder die Vereinigung aller Tags, also $\{FL,P\}$ als Tag.

        Hier gibt es einige Möglichkeiten, wie dieser Schritt genau aussieht.
        1. Version: Es wird nur der erste Tag vom Vaterknoten genutzt. Außerdem wird von den 
        Geschwisterknoten zuerst der Schnitt gebildet, und danach vom Ergebnis nochmal mit dem Vaterknoten zusammen.
        (Immer wenn der Schnitt leer ist, ist das Ergebnis die Vereinigung aller Tags, also $\{FL,P\}$. Auch im folgenden...)
        2. Version: Es wird nur der erste Tag vom Vaterknoten genutzt. Er wird zusammen mit den 
        Geschwistertags genommen und direkt ein Schnitt aller Mengen gebildet.
        3. Version: Es werden alle vorherigen Tags vom Vater genutzt und von diesen ein Schnitt gebildet.
        Das selbe gilt für die Geschwistertags. Und dann wird ein dritter Schnitt zwischend en Ergebnissen gebildet.
        4. Version: Es werden alle Tags genutzt und direkt in einem Schnitt zusammengenommen.

        Der Finale Schritt traversiert nochmal über den Baum und Bildet aus den zwei Tags pro Knoten einen finalen
        Tag, indem wieder der Schnitt der beiden Tags das Ergebnis ist.

@@

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Results}
  Ein großer Punkt in diesem Kapitel ist das Ergebnis der Untersuchung der Datenlage. Wie ist der 
    Stand? Welchen Einfluss hat das auf unser eigentliches Ergebnis? Was kann man dagegen tun? \\
  Ansonsten geht es in diesem Kapitel hauptsächlich um die tatsächliche Rekonstruktion der states.
    Dass heißt zum einen der der inneren Knoten und dadurch der Untersuchung von Origins und Losses
    und zum anderen der Vorhersage von unbekannten states von Leaf nodes.

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Tree}
    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Taxa}
      Die Untersuchung der Taxa der Eukaryoten hat das folgende ergeben:
      Es gibt 3 Kingdoms: Chloroplastida, Metazoa, Fungi. \\
      \todo{Hier fehlt nichts. Ref einfügen.} \\
      Phyla (53): \\
      Acanthocephala, Amoebozoa, Apicomplexa, Arthropoda, Ascomycota, Bacillariophyta, Basidiomycota, 
        Brachiopoda, Bryozoa, Chaetognatha, Chlorophyta, Chordata, Chromerida, Chytridiomycota, 
        Ciliophora, Cnidaria, Colponemidia, Ctenophora, Cycliophora, Echinodermata, Entoprocta, 
        Entorrhizomycota, Euglenida, Foraminifera, Gastrotricha, Glomeromycota, Gnathostomulida, 
        Haplosporida, Haptophyta, Hemichordata, Kinorhyncha, Loricifera, Microsporidia, Mollusca, 
        Myzostomida, Nematoda, Nematomorpha, Nemertea, Onychophora, Orthonectida, Phaeophyceae, 
        Picozoa, Placozoa, Platyhelminthes, Porifera, Priapulida, Rhodophyta, Rhombozoa, Rotifera, 
        Streptophyta, Tardigrada, Xanthophyceae \\
      Wobei von Streptophyta -> Anthocerotophyta, Marchantiophyta, Bryophyta, Tracheophyta als
        Phylum im Phylum gefunden und nicht einbezogen wurden und Magnoliophyta als Phylum in 
        Tracheophyta ebenfalls nicht. \\
      Es fällt auf, dass einige wichtige Phyla fehlen: \todo{Emanuel fragen}. \\
      Außerdem haben wir 195 Klassen und 924 Ordnungen feststellen können.

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Influence of different parameters}

>>>>

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{tag tree}
      \anmerkung{instead of 'tag tee' simulating states and transitions between them (Emanuel)} \\
      At this point we want one fully tagged tree, and one less tagged tree which looks like our real 
        data.\\
      Let's say the first specie (the root node) was free-living (start with a parasite without a host 
        makes no sence). For every transition from a node to his child, we take a random number from
        the father distribution. We decided that from the biological perspective a beta distribution
        reflects our transition probabilities best (see Figure 2.1 \todo{ref einfügen}). \\
      \anmerkung{why is that the case and why is that from a biological perspective? (Bernhard)} \\
      \anmerkungstext{I'd rather say that to ensure that the parameter of the binomial distribution is 
        restricted to the [0,1] interval, we model it... (Bernhard)} \\
      For example when our father node was free-living, then we take from the free-living beta
        distribution. Is the number under the threshold for beeing parasite, we get a change and tag
        the current node as parasite. Otherwise we tag it as free-living. \\
      With this procedure we traverse through the tree from the root to every leaf node. A part of
        this code you see here:
      \begin{lstlisting}[gobble=6]
        from numpy import random
        if father_tag == 0:
          # freeliving_distribution:
          new_random = random.beta(a=A_FL, b=B_FL)
        else:
          # parasite_distribution:
          new_random = random.beta(a=A_P, b=B_P)
        tag = 0         # $\rightarrow$ FL
        if new_random < percentage_parasites:
          tag = 1       # $\rightarrow$ P
      \end{lstlisting}
      \begin{figure}
        \caption{60\% Free-living - 40\% Parasites}
        \centering
          \includegraphics[width=0.5\textwidth]{Figures/40-60.png}
      \end{figure}
      \todo{Bessere Beschriftung, Plot neu erstellen! U.a. mit threshold} \\
      We save each tag with the associated node ID in a nodelist. \\
      \anmerkung{simulationg loss of information (1) states (2) topology/multifurcation (Emanuel)} \\
      The real tree has much less information, we have only information from some current species 
        (leaf nodes) and \todo{and probably negligible internal nodes}. \\
      To simulate our real tree we save for every node an empty placeholde except for some leaf nodes.
      There we save the states again. The amount of these unknown information is one parameter, which we
      got from our real tree. Or which we can change to \todo{...}
      \todo{Was hiervon gehört ind Methoden, was schon in Implementierung oder ganz woanders hin?}

@@

        Der Fitch Algorithmus ist aufgeteilt in drei Schritte, in welchen man jeweils durch den Baum traversiert.
        Schritt 1 beginnt von de Blättern aus, da sich dort zu Beginn die einzige Information befindet. 
        Für jeden Knoten gilt, wenn seine Kinder schon Information enthalten, dann bilde die Schnittmenge 
        der states und schreibe diese als Information in den aktuellen Knoten. Ist die Schnittmenge leer, 
        dann schreibe die Vereinigung aller möglichen states in den Knoten. Für alle Kinder, die noch keine 
        Information haben, führe diesen Schritt erst für diese aus.
        Schritt zwei geht von den Kindern der Wurzel bis zu den Vätern der Blätter. Jeder Knoten bekommt 
        einen zweiten Tag, der sich aus der Vereinigung des states des Vaterknoten und der Geschwisterknoten
        zusammensetzt. Ist diese leer, bekommt der Knoten wieder die Vereinigung aller states, also $\{FL,P\}$ als Tag.

        Hier gibt es einige Möglichkeiten, wie dieser Schritt genau aussieht.
        1. Version: Es wird nur der erste Tag vom Vaterknoten genutzt. Außerdem wird von den 
        Geschwisterknoten zuerst der Schnitt gebildet, und danach vom Ergebnis nochmal mit dem Vaterknoten zusammen.
        (Immer wenn der Schnitt leer ist, ist das Ergebnis die Vereinigung aller states, also $\{FL,P\}$. Auch im folgenden...)
        2. Version: Es wird nur der erste Tag vom Vaterknoten genutzt. Er wird zusammen mit den 
        Geschwisterstates genommen und direkt ein Schnitt aller Mengen gebildet.
        3. Version: Es werden alle vorherigen states vom Vater genutzt und von diesen ein Schnitt gebildet.
        Das selbe gilt für die Geschwisterstates. Und dann wird ein dritter Schnitt zwischend en Ergebnissen gebildet.
        4. Version: Es werden alle states genutzt und direkt in einem Schnitt zusammengenommen.

        Der Finale Schritt traversiert nochmal über den Baum und Bildet aus den zwei states pro Knoten einen finalen
        Tag, indem wieder der Schnitt der beiden states das Ergebnis ist.

@@

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Results}
  Ein großer Punkt in diesem Kapitel ist das Ergebnis der Untersuchung der Datenlage. Wie ist der 
    Stand? Welchen Einfluss hat das auf unser eigentliches Ergebnis? Was kann man dagegen tun? \\
  Ansonsten geht es in diesem Kapitel hauptsächlich um die tatsächliche Rekonstruktion der states.
    Dass heißt zum einen der der inneren Knoten und dadurch der Untersuchung von Origins und Losses
    und zum anderen der Vorhersage von unbekannten states von Leaf nodes. \\
  Und natürlich dem Ergebnis der Simulation.

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Metadata analysis}

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Taxa}
      Die Untersuchung der Taxa der Eukaryoten hat das folgende ergeben:
      Es gibt 3 Kingdoms: Chloroplastida, Metazoa, Fungi. \\
      \todo{Hier fehlt nichts. Ref einfügen.} \\
      Phyla (53): \\
      Acanthocephala, Amoebozoa, Apicomplexa, Arthropoda, Ascomycota, Bacillariophyta, Basidiomycota, 
        Brachiopoda, Bryozoa, Chaetognatha, Chlorophyta, Chordata, Chromerida, Chytridiomycota, 
        Ciliophora, Cnidaria, Colponemidia, Ctenophora, Cycliophora, Echinodermata, Entoprocta, 
        Entorrhizomycota, Euglenida, Foraminifera, Gastrotricha, Glomeromycota, Gnathostomulida, 
        Haplosporida, Haptophyta, Hemichordata, Kinorhyncha, Loricifera, Microsporidia, Mollusca, 
        Myzostomida, Nematoda, Nematomorpha, Nemertea, Onychophora, Orthonectida, Phaeophyceae, 
        Picozoa, Placozoa, Platyhelminthes, Porifera, Priapulida, Rhodophyta, Rhombozoa, Rotifera, 
        Streptophyta, Tardigrada, Xanthophyceae \\
      Wobei von Streptophyta -> Anthocerotophyta, Marchantiophyta, Bryophyta, Tracheophyta als
        Phylum im Phylum gefunden und nicht einbezogen wurden und Magnoliophyta als Phylum in 
        Tracheophyta ebenfalls nicht. \\
      Außerdem haben wir 195 Klassen und 924 Ordnungen feststellen können.

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Results of castor}

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Biological view}

      \todo{Castor replaces originaltags with finaltags. There are 82 originaltags != finaltag.} \\ \\

      We picked a few phyla to evaluate the results from the biological point of view. \\
      Table \ref{table:phylum leaf nodes states} shows some known phyla: Chordata, Nematoda, 
        Platyhelminthes and Apicomplexa \\
      As we know \todo{ref einfügen} the Chordata are full of free-living species and there are only a 
        few parasites. The algorithm reflects this. We started with 99.83\% free-living species and 
        predicted 99.94\% species as parasites \todo{(inklusive all known nodes)}. Only 0.06\% were 
        predicted as parasites. These few parasites are mostly breeding parasites (brood parasitism) and 
        some individual errors from the GloBI database. \\
      Same observation but with less free-livings is the Apicomplexa Phylum. Here we have only a few 
        free-livings \todo{ref einfügen}. And as we see, we had good start data and predicted 00.95\% as 
        parasites. \\
      For the Platyhelminthes the literature says that there are mostly all Platyhelminthes parasites
        \todo{ref einfügen}. But at the end we predicted 4.18\% as free-living. The class Seriata is the 
        reason for the most of free-livings in this phylum. These are partly free-living flatworms, so 
        the prediction looks right. \\
        \todotext{ Wiki: \\
        *The Seriata are an order of turbellarian flatworms.[1][2] \\
        They are found in both freshwater and marine environments, and also include a number of species 
        found in damp terrestrial conditions. Most are free-living, but the group includes the genus 
        Bdelloura, which lives comensally on the gills of horseshoe crabs. Seriatans are distinguished 
        from other related groups by the presence of a folded pharynx and of a number of diverticula 
        arising from the intestine. The intestine itself may be either simple or branched.[3]}

      With the Nematoda it looks very different.  In the Nematoda its much worse. Here speaks the 
        literature of \todo{x percent} free-living, but we found only 2.63\% of them. The problem at 
        this point, however, is obvious: The parasites have been much more studied and thus we start 
        with only 0.63\% free-living species. Against such a shifted data situation, the algorithm is 
        powerless.

      \begin{table}
        \begin{center}
          \begin{tabular}{ |l|r||r|r||r|r|r|r|r|r| }
            \hline
            & & \multicolumn{2}{c}{original states} & \multicolumn{6}{c}{final states} \\
            Phylum & \# nodes & (FL) & (P) 
              & 0 (FL) & 0.4 & 0.5 & 0.667 & 0.75 & 1 (P) \\
            \hline \hline
            Chordata & 91785 & 10451 & 18 
              & 91734 & 0 & 0 & 0 & 0 & 51 \\
            & & 99.83\% & 0.49\%
              & 99.94\% & & & & & 0.06\% \\ \hline
            Nematoda & 30127 & 21 & 3289 
              & 791 & 0 & 1017 & 0 & 0 & 28319 \\
            & & 0.63\% & 99.37\%
              & 2.63\% & & 3.38\% & & & 94\% \\ \hline
            Platyhelminthes & 22683 & 7 & 7086 
              & 949 & 0 & 151 & 0 & 0 & 21583 \\
            & & 0.1\% & 99.9\%
              & 4.18\% & & 0.67\% & & & 95.15\% \\ \hline
            Apicomplexa & 1863 & 1 & 255 
              & 1 & 0 & 0 & 0 & 0 & 1862 \\
            & & 0.39\% & 99.61\%
              & 0.05\% & & & & & 99.95\% \\
            \hline \hline
            Anthropoda & 1198981 & 18912 & 11141 
              & 1099509 & 1313 & 22478 & 4176 & 1665 & 70223 \\
            & & 62.93\% & 37.07\%
              & 91.7\% & 0.11\% & 1.87\% & 0.35\% & 0.14\% & 5.86\% \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{Phylum (leaf nodes)}
        \label{table:phylum leaf nodes states} 
      \end{table}

      \anmerkungstext{Das sind schonmal vier große Kontraste, wenn dann noch Zeit bleibt, die 
        schwirigen... Arthropoden, Fungi, Pflanzen... (Emanuel)} \\

      \begin{table}
        \begin{center}
          \begin{tabular}{ |l|r||r|r||r|r|r|r|r|r|r|r| }
            \hline
            & & \multicolumn{2}{c}{original states} & \multicolumn{8}{c}{final states} \\
            Kingdom & \# nodes & (FL) & (P) 
              & 0 (FL) & 0.25 & 0.333 & 0.4 & 0.667 & 0.75 & 1 (P) \\
            \hline \hline
            none & 84456 & 45 & 529 
              & 15035 & 243 & 25910 & 0 & 8764 & 6183 & 0 & 28140 \\
            Fungi & 324105 & 577 & 2983
              & 39088 & 0 & 0 & 0 & 5858 & 0 & 0 & 274803 \\
            Chloroplastida & 460457 & 3519 & 77
              & 454211 & 0 & 0 & 0 & 4688 & 0 & 0 & 1558 \\
            Metazoa & 1670956 & 30758 & 22373
              & 1485749 & 0 & 0 & 1313 & 29002 & 5102 & 1957 & 147833 \\
            \hline  
          \end{tabular}
        \end{center}
        \caption{Kingdom (inkl internal nodes)}
      \end{table}

      \begin{table}
        \begin{center}
          \begin{tabular}{ |l|r||r|r||r|r|r|r|r|r| }
            \hline
            & & \multicolumn{2}{c}{original states} & \multicolumn{6}{c}{final states} \\
            Phylum & \# nodes & (FL) & (P) 
              & 0 (FL) & 0.4 & 0.5 & 0.667 & 0.75 & 1 (P) \\
            \hline \hline
            Chordata & 122546 & 10451 & 18 
              & 122473 & 0 & 0 & 0 & 0 & 73 \\
            Nematoda & 33564 & 21 & 3289 
              & 846 & 0 & 1133 & 0 & 0 & 31585 \\
            Platyhelminthes & 27142 & 7 & 7086 
              & 1010 & 0 & 175 & 0 & 0 & 25957 \\
            Apicomplexa & 2102 & 1 & 255 
              & 1 & 0 & 0 & 0 & 0 & 2101 \\
            \hline
            Anthropoda & 1319460 & 18912 & 11141 
              & 1207204 & 1313 & 25499 & 4852 & 1957 & 78635 \\
            \hline
          \end{tabular}
        \end{center}
        \caption{Phylum (inkl internal nodes)}
      \end{table}

      \begin{table}
        \begin{center}
          \begin{tabular}{ |l|r||r|r||r|r|r|r|r|r|r|r| }
            \hline
            & & \multicolumn{2}{c}{original states} & \multicolumn{8}{c}{final states} \\
            Kingdom & \# nodes & (FL) & (P) 
              & 0 (FL) & 0.25 & 0.333 & 0.4 & 0.5 & 0.667 & 0.75 & 1 (P) \\
            \hline \hline
            none & 75446 & 45 & 529 
              & 13426 & 220 & 24082 & 0 & 7792 & 5302 & 0 & 24493 \\
            Fungi & 31457 & 577 & 2983
              & 38520 & 0 & 0 & 0 & 5723 & 0 & 0 & 266463 \\
            Chloroplastida & 416478 & 3519 & 77
              & 410795 & 0 & 0 & 0 & 4182 & 0 & 0 & 1501 \\
            Metazoa & 1491012 & 30758 & 22373
              & 1328135 & 0 & 0 & 930 & 25535 & 4423 & 1665 & 130324 \\
            \hline  
          \end{tabular}
        \end{center}
        \caption{Kingdom (leaf nodes)}
      \end{table}

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Results of simulation}

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Influence of different parameters}
